### 2019-05-24 revtools/analysis.Rmd

---
title: "Untitled"
author: "Aaron Conway"
date: "29/04/2019"
output: html_document
---

```{r}
library(revtools)
library(tidyverse)
library(vroom)
```

# Search pubmed with phrase "zero heat flux"

```{r}
data <- read_bibliography("citations.bib")
```

# Screen abstracts

*It is essential that data is saved as .csv before exiting webpaage*

```{r}
screen_abstracts(data)
```

# Load in results from reviewers

```{r message=FALSE}
reviewer_1 <- read_csv("reviewer_1.csv") %>% 
  mutate(reviewer_one = selected)

reviewer_2 <- read_csv("reviewer_2.csv") %>% 
  mutate(reviewer_two = selected)

abstract_review <- right_join(reviewer_1, reviewer_2, by = "label")
abstract_review <- abstract_review %>% 
  select(label, title.x, abstract.x, doi.x, author.x, reviewer_one, reviewer_two)
names <- c("label", "title", "abstract", "doi", "author", "reviewer_one", "reviewer_two")
colnames(abstract_review) <- names
```

# Compare results between reviewers

```{r}
# Identify inclusions (both reviewers selected include)

inclusions <- abstract_review %>% 
  filter(reviewer_one == "selected" & reviewer_two == "selected")

# Identify exclusions (both reviewers selected include)

exclusions <- abstract_review %>% 
   filter(reviewer_one == "excluded" & reviewer_two == "excluded")

# Identify conflicts (diagreement among reviewers)

conflicts <- abstract_review %>% 
  filter((reviewer_one == "excluded" & reviewer_two == "selected") | (reviewer_one == "selected" & reviewer_two == "excluded"))

# Third reviewer to resolve conflicts

screen_abstracts(as.data.frame(conflicts))

# Add inclusions and exclusions after resolution of conflicts

reviewer_3 <- read_csv("reviewer_2.csv")

inclusions_3 <- reviewer_3 %>% 
  filter(selected == "selected")

exclusions_3 <- reviewer_3 %>% 
  filter(selected == "excluded")

```

# Full text screening

```{r}
library(BibScan)
```

```{r}
library(crminer)
x <- "6c00c459-ddeb8bce-425652b1-aa68cd58"
Sys.setenv(CROSSREF_TDM=x)
e <- "aaron.conway@utoronto.ca"
Sys.setenv(CROSSREF_EMAIL=e)
```

```{r}
indir <- here::here("full-texts", "savedrecs.bib")
outdir <- here::here('full-texts')
article_pdf_download(indir, outdir)
```

```{r}
df_citations <- bibfile_reader(indir, bib_format)
```