### 2019-08-12 zhf_manuscript/flow_nomnoml.Rmd
study inclusion flow chart

---
title: "flow diagram"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(nomnoml)
```
```{nomnoml}
#import: my-common-styles.nomnoml
#stroke: black
#.blob: fill=#DBDBDB visual=italic
#direction: down
#arrowSize: 0.6
#bendSize: 0.3
#fillArrows: true

  [129 references imported for screening]->[94 studies screened against title and abstract]

  [129 references imported for screening]-->[<blob> 35 duplicates removed]

  [94 studies screened against title and abstract]->[23 studies assessed for full-text eligibility]
  [94 studies screened against title and abstract]-->[<blob> 71 studies excluded]

[23 studies assessed for full-text eligibility] -> [17 studies included]

 [23 studies assessed for full-text eligibility] --> [<blob> 7 studies excluded | 4 Wrong intervention

                                                      1 Wrong comparator

                                                      1 Wrong study design]
```