### 2019-08-27 zhf_manuscript/zhf_manuscript.tex

% !TeX program = pdfLaTeX
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

%% load any required packages here

\begin{document}

\title{Accuracy and precision of zero-heat-flux temperature monitoring: a
systematic review and meta- analysis \thanks{Grants or other notes about the article that should go on the front page
should be placed here. General acknowledgments should be placed at the
end of the article.} }

    \titlerunning{Short form of title (if too long for head)}

\author{  Aaron Conway \and  Âuthóř 2 \and  }

    \authorrunning{ Conway, }

\institute{
        Aaron Conway \at
     Peter Munk Cardiac Centre, UHN \& Lawrence S. Bloomberg Faculty of
 Nursing, University of Toronto \\
     \email{\href{mailto:aaron.conway@utoronto.ca}{\nolinkurl{aaron.conway@utoronto.ca}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Âuthóř 2 \at
     Department of ZZZ, University of WWW \\
     \email{\href{mailto:djf@wef}{\nolinkurl{djf@wef}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
The text of your abstract. 150 -- 250 words.
\\
\keywords{
        key \and
        dictionary \and
        word \and
    }

    \subclass{
                    MSC code 1 \and
                    MSC code 2 \and
            }

\end{abstract}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Core temperature monitoring is an important intervention within
perioperative and intensive care settings\((ref)\). Thermoregulatory
dysfunction is commonly associated with the induction of anesthesia and
can lead to adverse outcomes including cardiac dysrhythmias, altered
hemostasis and increased risk of surgical site infection (Frank et al.
1997; Kurz, Sessler, and Lenhardt 1996; Michelson et al. 1994; Rohrer
and Natale 1992). The pulmonary artery catheter is the reference
standard for continuous core temperature monitoring\((ref)\), however,
the invasive nature of this method renders an increased risk of
bloodstream infections and damage to surrounding tissue(Hadian and
Pinsky 2006). Although clinically accurate relative to pulmonary artery
thermometry, surrogate measures of core temperature via esophageal,
rectal, and bladder sites remain mildly invasive interventions\((ref)\).

Zero-heat-flux (ZHF) thermometry is an alternative non-invasive method
that allows for continuous monitoring of core temperature. Originally
developed in the 1970s, ZHF technology was recently implemented in the
3M SpotOn (3M, St Paul, MN), a single-use, disposable sensor \((ref)\).
In practice, the ZHF sensor, such as the 3M SpotOn, is placed on the
lateral surface of the forehead and initially warmed to equilibrate the
temperature of the skin surface to the underlaying core
tissues\((ref)\). Equipped with a thermal insulator, the ZHF sensor
eliminates heat loss to the environment to allow for changes in core
temperature to be directly reflected by a change in skin surface
temperature \((ref)\).

The agreement between ZHF and established core temperature measures has
been investigated in several studies conducted within the clinical
setting, each with varied results. Appraisal of these studies and
synthesis of the results would aid
\textbackslash{}\clinical {decision-making regarding the appropriate circumstances in which ZHF monitoring may be used.}

We aimed to determine if the ZHF sensor has clinically acceptable
accuracy and precision compared with established core temperature tools.
Accuracy is defined as the average difference between the ZHF sensor and
established core temperature instrument and precisions as the variance
(typically reported as the SD) in the differences.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

A systematic review was conducted in accordance with a predetermined
protocol (PROSPERO trial registration number:
\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#).

\hypertarget{data-sources-and-searches}{%
\subsection{Data sources and searches}\label{data-sources-and-searches}}

Published studies were found by searching Medline and EMBASE from
January 2000 to July 2019, \textgreater{}as well as the reference lists
of articles identified to be relevant to the review. This search
strategy is an efficient approach for systematic reviews of diagnostic
test accuracy studies (Preston et al. 2015). Unpublished and ongoing
studies were located by searching the International Clinical Trials
Platform.

Published conference abstracts were included if there was enough
information reported to appraise the quality of the study. There were no
language restrictions applied during the search. \textgreater{}The
Cochrane-recommended search strategy combining terms for the `target
condition' and `index test' was used (Mann and Gilbody 2012).

The search strategies used for each data base can be located in
\href{}{file 1}.

The selection of studies was undertaken by two independent reviewers.
Studies were included if reported measures from a ZHF sensor and another
established core temperature tool coincided. Studies involving a case
control design were excluded due to potential for overestimation of the
intervention performance. Studies were excluded if conducted on
non-human subjects or outside of a clinical healthcare setting. Studies
were excluded if conducted earlier than the year 2000 as prior studies
evaluated outdated technology that is no longer relevant to current
clinical practice.

\hypertarget{data-extraction-and-quality-assessment}{%
\subsection{Data extraction and quality
assessment}\label{data-extraction-and-quality-assessment}}

\begin{quote}
Information was extracted regarding study characteristics(author, year
of publication, country, design, sample size, clinical setting, numbers
studied and analyses for each outcome), population characteristics
(inclusion and exclusion criteria) and ZHF characteristics (timing and
methods of measurements, and methods of calibration).
\end{quote}

The outcomes that were extracted included the mean bias (eg, accuracy)
and variance (eg, SD, precision) in temperature between the ZHF sensor
and established core temperature measurement tool.
\textgreater{}Information was extracted about how repeated measurements
were handled: (1) analysed each pair of data separately; (2) treated
each pair of data as independent; or (3) used either analysis of
variance or a random effects model as a way to control for the dependent
nature of the repeated measures data (Myles and Cui 2007).

Using the revised Quality Assessment of Diagnostic Accuracy Studies
(QUADAS-2), each reviewer independently assessed the risk of bias for
the included studies (Whiting et al. 2011). \textgreater{}Reviewers used
guiding questions to rate the risk of bias for patient selection,

conduct of the ZHF measurements, conduct of the established core
temperature measurement tools, and timing and flow (eg, timing of ZHF
and established core temperature measurements, dropouts) as ``high'',
``low'' or ``unclear''. We worked to minimize the risk of publication
bias by conducting a comprehensive search of multiple databases as well
as an international clinical trial registry (Glasziou et al. 2001).
\textgreater{}Statistical analyses to detect reporting bias were not
conducted due to lack of validated methods (Begg 2005). Although some
meta-analyses of method comparison studies have used tests for detecting
funnel plot asymmetry (Niven et al. 2015), simulations have revealed
that such tests will result in publication bias being incorrectly
identified too often (Deeks, Macaskill, and Irwig 2005).

In order to rate the quality of evidence, we applied the Grading Quality
of Evidence and Strength of Recommendation methodology (Schünemann et
al. 2008). Evidence was downgraded in accordance with study limitations,
inconsistency and imprecision. There were no circumstances in which
evidence was downgraded for indirectness as this systematic review only
included relevant studies. Although the possibility of publication bias
was not excluded, this bias was not formally assessed as it was not
considered sufficient enough to reason downgrading the quality of
evidence.

\hypertarget{data-synthesis-and-analysis}{%
\subsection{Data synthesis and
analysis}\label{data-synthesis-and-analysis}}

The objective
\textgreater{}\(for the meta-analysis was to estimate the population limits of agreement between\)

temperature measurements by the ZHF sensor and established core
temperature measurement tools. \textgreater{}The framework for
meta-analysis of Bland-Altman method comparison studies based on limits
of agreement (LoA) approach was used (Tipton and Shuster 2017).

This method was selected because it parallels the approach used in
primary Bland-Altman studies, whereby an estimate is generated for the
pooled LoAs in the population (not just in the samples studied). The
`population LoA' is more broad than the LoA commonly reported in the
meta-analyses of Bland-Altman studies (Tipton and Shuster 2017). In
these instances, \textgreater{}the pooled LoAs are calculated using
\(\delta\pm2\sqrt{\sigma2+\tau2}\), where \(\delta\) is the average bias
across studies, \(\sigma2\) is the average within-study variation in
differences and \(\tau2\) is the variation in bias across studies.

Estimations of \(\delta\) and \(\sigma2\) were made using a weighted
least-squares model (similar to a random-effects approach), with the
associated SEs estimated using robust variance estimation (RVE). RVE was
used alternatively to model-based SEs as several studies included in the
systematic review used repeated-measures designs without accommodating
for the correlation between measurements (Hedges, Tipton, and Johnson
2010; Tanner-Smith, Tipton, and Polanin 2016; Tipton 2015). We used the
\textgreater{}method-of-moments estimator from ref 16 for the \(\tau2\)
parameter.

In accordance with ref 12, (1) measures of uncertainty were included
\textgreater{}when interpreting the LoA estimates by calculating the
outer \(95\%\) CIs for pooled LoA;

and (2) repeated measurements were adjusted if improperly adjusted for
in \textgreater{}individual studies (by using weights proportional to
the number of samples, not the total number of measurements).

The formulas for the above calculations from ref 12 can be found in
\href{}{file 1}. The R statistical program was used to conduct all
analyses (Team 2017). All data and R code (provided in ref 12) used in
the meta-analyses can be located \href{https://doi.org}{here}.

Prior to conducting the meta-analyses, the results from each study were
converted into a standard format, with bias meaning \(ECT - ZHF\)
measured in \(^\circ C\). In two of the studies, the results were
reported for two separate groups of participants, so these were treated
in the meta-analyses as separate `studies'. \textgreater{}Other studies
reported separate results for analyses conducted using different TcCO2
device types or sensor locations performed on the same patients. Only
the result with the largest number of paired measurements between PaCO2
and TcCO2 was selected for inclusion in the main analysis, with others
included in subgroup meta-analyses where appropriate. For the studies
that reported results for patients while receiving both two-lung and
one-lung ventilation during thoracic surgery, we used the result for
two-lung ventilation in the main analysis. The conventionally cited
clinically acceptable agreement between

ECT and ZHF sensors is 0.5 \(^\circ C\)(ref).
\textgreater{}\(It was deemed that outer confidence bounds for 95\% LoA between ECT and ZHF (termed as ‘population limits of agreement’) outside of these bounds would\)

be clinically unacceptable. Sensitivity analysis for the primary
meta-analysis was performed based on risk of bias, whereby `unclear risk
of bias' was treated as `high risk' and `high risk of bias' studies were
excluded from analyses. \textgreater{}As clinicians would be interested
in the accuracy of

ZHF relative to thermometer devices they use, and within the clinical
setting in which they use it we conducted subgroup analyses according to
the comparator device used (eg, core, peripheral, nasopharangeal) and
clinical setting (eg, intraoperative, intensuve care unit).

\hypertarget{results}{%
\section{Results}\label{results}}

Sixteen studies were eligible for inclusion in this systematic review.
\textgreater{}The characteristics of each study are in online
supplementary \href{link\%20to\%20file\%201}{file 1}.

Using these studies, we conducted 6 different meta-analyses. The first
group compared ZHF to core (eg, arterial, bladder, esophageal, rectal)
temperature measurements and included 22 (73\%) comparisons from 14 (\%)
studies, and enrolled 607 participants with 179,821 paired measurements.
\textgreater{}The second group compared ZHF to core temperature
measurements in low risk bias studies studies that scored a low risk of
bias in all categories) and included 10 (\%) comparisons from 5 (\%)
studies, and enrolled 273 participants with 104,294 paired measurements.

The third group compared ZHF to core temperature measurements in ICU
patients and included 7 (23\%) comparisons from 5 (31\%) studies, and
enrolled 246 participants with 155,598 paired measurements. The fourth
group compared ZHF to core temperature measurement in intraoperative
patients and included 16 (53\%) comparisons from 10 (63\%) studies, and
enrolled 464 participants with 24,223 paired measurements. The fifth
group compared ZHF to nasopharyngeal temperature measurements and
included 4 (13\%) comparisons from 4 (25\%) studies, and enrolled 297
participants with 109,819 paired measurements. The sixth group compared
ZHF to sublingual temperature measurement and included 2 (7\%)
comparisons from 2 (13\%) studies, and enrolled 107 participants with
22731 paired measurements. The studies included in this systematic
review evaluated a ZHF temperature monitoring system manufactured by 3M.
Previously known as the SpotOn Temperature Monitoring System, the 3M ZHF
device is now referred to as the Bair Hugger Temperature Monitoring
System. All comparisons reported adherence to the ZHF device
manufacturer instructions and placed the sensor on the forehead of the
participants. High risk of bias was associated with patient selection
for 13 (43\%) comparisons from 9 (56\%) studies, conduct of ZHF and
comparator measurements in 10(33\%) comparisons from 8 (50\%) studies
and 12 (40\%) comparisons 9 (30\%) studies, respectively (mostly due to
ZHF measurements being taken with knowledge of the comparator
measurements and vice versa) and participant flow for 13 (43\%) studies.
In 19 (50\%) comparisons from 8 (50\%) studies, the authors had declared
a conflict of interest or receipt of funding or equipment from the
manufacturer of the ZHF device under evaluation.

Table 1
\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}
\textbar{}Table 1 Results of meta-analysis of agreement between
transcutaneous and arterial carbon dioxide\textbar{}
\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}
\textbar{}---\textbar{}Studies\textbar{}N\textbar{}n\textbar{}Bias\textbar{}SD\textbar{}tau2\textbar{}LoA\_L\textbar{}LoA\_U\textbar{}CI\_L\textbar{}CI\_U\textbar{}
\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}
\textbar{}Subset\textbar{}Stu-content\textbar{}N-content\textbar{}n-content\textbar{}Bias-content\textbar{}SD-content\textbar{}tau2-content\textbar{}LoAL-content\textbar{}LoAU-content\textbar{}CIL-content\textbar{}CIU-content\textbar{}
\textbar{}Subset\textbar{}Stu-content\textbar{}N-content\textbar{}n-content\textbar{}Bias-content\textbar{}SD-content\textbar{}tau2-content\textbar{}LoAL-content\textbar{}LoAU-content\textbar{}CIL-content\textbar{}CIU-content\textbar{}

Table 2 \textbar{}Outcome\textbar{}Summary accuracy and precision
(C)\textbar{}Participants\textbar{}Study
characteristics\textbar{}Quality\textbar{}Implications\textbar{}
\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}---\textbar{}
\textbar{}Out-content\textbar{}Sum-content\textbar{}Par-content\textbar{}Char-content\textbar{}Qual-content\textbar{}Impl-content\textbar{}
\textbar{}Out-content\textbar{}Sum-content\textbar{}Par-content\textbar{}Char-content\textbar{}Qual-content\textbar{}Impl-content\textbar{}
\textbar{}Out-content\textbar{}Sum-content\textbar{}Par-content\textbar{}Char-content\textbar{}Qual-content\textbar{}Impl-content\textbar{}
\textbar{}Out-content\textbar{}Sum-content\textbar{}Par-content\textbar{}Char-content\textbar{}Qual-content\textbar{}Impl-content\textbar{}

Table 3 \textbar{} test1 \textbar{} test2 \textbar{} test3 \textbar{}
test4 \textbar{} test5 \textbar{}
\textbar{}-------\textbar{}-------\textbar{}-------\textbar{}-------\textbar{}-------\textbar{}
\textbar{} a \textbar{} b \textbar{} c \textbar{} d \textbar{} e
\textbar{} \textbar{} a \textbar{} b \textbar{} c \textbar{} d
\textbar{} e \textbar{} \textbar{} a \textbar{} b \textbar{} c
\textbar{} d \textbar{} e \textbar{}

\# simple table creation here \textbar{} Tables \textbar{} Are
\textbar{} Cool \textbar{}
\textbar{}---------------\textbar{}:-------------:\textbar{}------:\textbar{}
\textbar{} col 3 is \textbar{} right-aligned \textbar{} \$1600
\textbar{} \textbar{} col 2 is \textbar{} centered \textbar{} \$12
\textbar{} \textbar{} zebra stripes \textbar{} are neat \textbar{} \$1
\textbar{}

\hypertarget{study-selection-and-description}{%
\subsection{Study selection and
description}\label{study-selection-and-description}}

\hypertarget{agreement-between-zhf-and-core-temperature-measurements}{%
\subsection{Agreement between ZHF and core temperature
measurements}\label{agreement-between-zhf-and-core-temperature-measurements}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-begg2005systematic}{}%
Begg, Colin B. 2005. ``Systematic Reviews of Diagnostic Accuracy Studies
Require Study by Study Examination: First for Heterogeneity, and Then
for Sources of Heterogeneity.'' \emph{Journal of Clinical Epidemiology}
58 (9). Elsevier Limited: 865.

\leavevmode\hypertarget{ref-deeks2005performance}{}%
Deeks, Jonathan J, Petra Macaskill, and Les Irwig. 2005. ``The
Performance of Tests of Publication Bias and Other Sample Size Effects
in Systematic Reviews of Diagnostic Test Accuracy Was Assessed.''
\emph{Journal of Clinical Epidemiology} 58 (9). Elsevier: 882--93.

\leavevmode\hypertarget{ref-frank1997perioperative}{}%
Frank, Steven M, Lee A Fleisher, Michael J Breslow, Michael S Higgins,
Krista F Olson, Susan Kelly, and Charles Beattie. 1997. ``Perioperative
Maintenance of Normothermia Reduces the Incidence of Morbid Cardiac
Events: A Randomized Clinical Trial.'' \emph{Jama} 277 (14). American
Medical Association: 1127--34.

\leavevmode\hypertarget{ref-glasziou2001systematic}{}%
Glasziou, Paul, Les Irwig, Chris Bain, and Graham Colditz. 2001.
\emph{Systematic Reviews in Health Care: A Practical Guide}. Cambridge
University Press.

\leavevmode\hypertarget{ref-hadian2006evidence}{}%
Hadian, Mehrnaz, and Michael R Pinsky. 2006. ``Evidence-Based Review of
the Use of the Pulmonary Artery Catheter: Impact Data and
Complications.'' \emph{Critical Care} 10 (3). BioMed Central: S8.

\leavevmode\hypertarget{ref-hedges2010robust}{}%
Hedges, Larry V, Elizabeth Tipton, and Matthew C Johnson. 2010. ``Robust
Variance Estimation in Meta-Regression with Dependent Effect Size
Estimates.'' \emph{Research Synthesis Methods} 1 (1). Wiley Online
Library: 39--65.

\leavevmode\hypertarget{ref-kurz1996perioperative}{}%
Kurz, Andrea, Daniel I Sessler, and Rainer Lenhardt. 1996.
``Perioperative Normothermia to Reduce the Incidence of Surgical-Wound
Infection and Shorten Hospitalization.'' \emph{New England Journal of
Medicine} 334 (19). Mass Medical Soc: 1209--16.

\leavevmode\hypertarget{ref-mann2012should}{}%
Mann, Rachel, and Simon M Gilbody. 2012. ``Should Methodological Filters
for Diagnostic Test Accuracy Studies Be Used in Systematic Reviews of
Psychometric Instruments? A Case Study Involving Screening for Postnatal
Depression.'' \emph{Systematic Reviews} 1 (1). BioMed Central: 9.

\leavevmode\hypertarget{ref-michelson1994reversible}{}%
Michelson, Alan D, Hollace MacGregor, Marc R Barnard, Anita S Kestin,
Michael J Rohrer, and C Robert Valeri. 1994. ``Reversible Inhibition of
Human Platelet Activation by Hypothermia in Vivo and in Vitro.''
\emph{Thrombosis and Haemostasis} 72 (05). Schattauer GmbH: 633--40.

\leavevmode\hypertarget{ref-myles2007using}{}%
Myles, PS, and James Cui. 2007. ``I. Using the Bland--Altman Method to
Measure Agreement with Repeated Measures.'' Oxford University Press.

\leavevmode\hypertarget{ref-niven2015accuracy}{}%
Niven, Daniel J, Jonathan E Gaudet, Kevin B Laupland, Kelly J Mrklas,
Derek J Roberts, and Henry Thomas Stelfox. 2015. ``Accuracy of
Peripheral Thermometers for Estimating Temperature: A Systematic Review
and Meta-Analysis.'' \emph{Annals of Internal Medicine} 163 (10).
American College of Physicians: 768--77.

\leavevmode\hypertarget{ref-preston2015improving}{}%
Preston, Louise, Christopher Carroll, Paolo Gardois, Suzy Paisley, and
Eva Kaltenthaler. 2015. ``Improving Search Efficiency for Systematic
Reviews of Diagnostic Test Accuracy: An Exploratory Study to Assess the
Viability of Limiting to Medline, Embase and Reference Checking.''
\emph{Systematic Reviews} 4 (1). BioMed Central: 82.

\leavevmode\hypertarget{ref-rohrer1992effect}{}%
Rohrer, Michael J, and ANITA M Natale. 1992. ``Effect of Hypothermia on
the Coagulation Cascade.'' \emph{Critical Care Medicine} 20 (10):
1402--5.

\leavevmode\hypertarget{ref-schunemann2008grading}{}%
Schünemann, Holger J, Andrew D Oxman, Jan Brozek, Paul Glasziou, Roman
Jaeschke, Gunn E Vist, John W Williams, et al. 2008. ``Grading Quality
of Evidence and Strength of Recommendations for Diagnostic Tests and
Strategies.'' \emph{Bmj} 336 (7653). British Medical Journal Publishing
Group: 1106--10.

\leavevmode\hypertarget{ref-tanner2016handling}{}%
Tanner-Smith, Emily E, Elizabeth Tipton, and Joshua R Polanin. 2016.
``Handling Complex Meta-Analytic Data Structures Using Robust Variance
Estimates: A Tutorial in R.'' \emph{Journal of Developmental and
Life-Course Criminology} 2 (1). Springer: 85--112.

\leavevmode\hypertarget{ref-team2017r}{}%
Team, R Core. 2017. ``R Core Team (2017). R: A Language and Environment
for Statistical Computing.'' \emph{R Found. Stat. Comput. Vienna,
Austria. URL Http://Www. R-Project. Org/., Page R Foundation for
Statistical Computing}.

\leavevmode\hypertarget{ref-tipton2015small}{}%
Tipton, Elizabeth. 2015. ``Small Sample Adjustments for Robust Variance
Estimation with Meta-Regression.'' \emph{Psychological Methods} 20 (3).
American Psychological Association: 375.

\leavevmode\hypertarget{ref-tipton2017framework}{}%
Tipton, Elizabeth, and Jonathan Shuster. 2017. ``A Framework for the
Meta-Analysis of Bland--Altman Studies Based on a Limits of Agreement
Approach.'' \emph{Statistics in Medicine} 36 (23). Wiley Online Library:
3621--35.

\leavevmode\hypertarget{ref-whiting2011quadas}{}%
Whiting, Penny F, Anne WS Rutjes, Marie E Westwood, Susan Mallett,
Jonathan J Deeks, Johannes B Reitsma, Mariska MG Leeflang, Jonathan AC
Sterne, and Patrick MM Bossuyt. 2011. ``QUADAS-2: A Revised Tool for the
Quality Assessment of Diagnostic Accuracy Studies.'' \emph{Annals of
Internal Medicine} 155 (8). Am Coll Physicians: 529--36.

\bibliographystyle{spbasic}
\bibliography{bibliography.bib}

\end{document}