### 2019-08-29 zhf_manuscript/zhf_manuscript.tex
prisma diagram

% !TeX program = pdfLaTeX
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

%% load any required packages here

\begin{document}

\title{Accuracy and precision of zero-heat-flux temperature monitoring: a
systematic review and meta- analysis \thanks{Grants or other notes about the article that should go on the front page
should be placed here. General acknowledgments should be placed at the
end of the article.} }

    \titlerunning{Short form of title (if too long for head)}

\author{  Aaron Conway \and  Âuthóř 2 \and  }

    \authorrunning{ Conway, }

\institute{
        Aaron Conway \at
     Peter Munk Cardiac Centre, UHN \& Lawrence S. Bloomberg Faculty of
 Nursing, University of Toronto \\
     \email{\href{mailto:aaron.conway@utoronto.ca}{\nolinkurl{aaron.conway@utoronto.ca}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Âuthóř 2 \at
     Department of ZZZ, University of WWW \\
     \email{\href{mailto:djf@wef}{\nolinkurl{djf@wef}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
The text of your abstract. 150 -- 250 words.
\\
\keywords{
        key \and
        dictionary \and
        word \and
    }

    \subclass{
                    MSC code 1 \and
                    MSC code 2 \and
            }

\end{abstract}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Core temperature monitoring is an important intervention within
perioperative and intensive care settings. Thermoregulatory dysfunction
is commonly associated with the induction of anesthesia and can lead to
adverse outcomes including cardiac dysrhythmias, altered hemostasis and
increased risk of surgical site infection (Frank et al. 1997; Kurz,
Sessler, and Lenhardt 1996; Michelson et al. 1994; Rohrer and Natale
1992). The pulmonary artery catheter is the reference standard for
continuous core temperature
monitoring\((ref - trouble finding a statement regarding this in the literature)\),
however, the invasive nature of this method renders an increased risk of
bloodstream infections and damage to surrounding tissue(Hadian and
Pinsky 2006). Although clinically accurate relative to pulmonary artery
temperature measurements, surrogate measures of core temperature at
esophageal, rectal, and bladder locations remain mildly invasive
interventions\((ref)\).

Zero-heat-flux (ZHF) thermometry is an alternative non-invasive method
that allows for continuous monitoring of core temperature. Originally
developed in the 1970s, ZHF technology was recently implemented in the
3M SpotOn (3M, St Paul, MN), a single-use, disposable sensor \((ref)\).
In practice, the ZHF sensor, such as the 3M SpotOn, is placed on the
lateral surface of the forehead and initially warmed to equilibrate the
temperature of the skin surface to the underlaying core
tissues\((ref)\). Equipped with a thermal insulator, the ZHF sensor
eliminates heat loss to the environment to allow for changes in core
temperature to be directly reflected by a change in skin surface
temperature \((ref)\).

The agreement between ZHF and established core temperature devices has
been investigated in several studies conducted within the clinical
setting, each with varied results. Appraisal of these studies and
synthesis of the results would aid clinicians in deciding the
appropriate circumstances in which ZHF may be used.

We aimed to determine if ZHF has clinically acceptable accuracy and
precision relative to established core temperature measurement devices.
Accuracy is defined as the average difference between temperature
measurements from the ZHF and comparator device and precisions as the
variance (typically reported as the SD) in the differences.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

A systematic review was conducted in accordance with a predetermined
protocol (PROSPERO trial registration number:
\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#\#).

\hypertarget{data-sources-and-searches}{%
\subsection{Data sources and searches}\label{data-sources-and-searches}}

Published studies were found by searching Medline and EMBASE from
January 2000 to July 2019, \textgreater{}as well as the reference lists
of articles identified to be relevant to the review. This search
strategy is an efficient approach for systematic reviews of diagnostic
test accuracy studies (Preston et al. 2015). Unpublished and ongoing
studies were located by searching the International Clinical Trials
Platform.

Published conference abstracts were included if there was enough
information reported to appraise the quality of the study. There were no
language restrictions applied during the search. \textgreater{}The
Cochrane-recommended search strategy combining terms for the `target
condition' and `index test' was used (Mann and Gilbody 2012).

The search strategies used for each data base can be located in
\href{}{file 1}.

The selection of studies was undertaken by two independent reviewers.
Studies were included if reported measures from a ZHF and comparator
device coincided. Studies involving a case control design were excluded
due to potential for overestimation of the intervention performance.
Studies were excluded if conducted on non-human subjects or outside of a
clinical healthcare setting. Studies were excluded if conducted earlier
than the year 2000 as prior studies evaluated outdated technology that
is no longer relevant to current clinical practice.

\hypertarget{data-extraction-and-quality-assessment}{%
\subsection{Data extraction and quality
assessment}\label{data-extraction-and-quality-assessment}}

Information was extracted regarding study characteristics (author, year
of publication, country, design, sample size, clinical setting, numbers
studied and analyses for each outcome), population characteristics
(inclusion and exclusion criteria) and ZHF characteristics (placement of
sensor, timing and methods of measurements). The outcomes that were
extracted included the mean bias (eg, accuracy) and variance (eg, SD,
precision) in temperature measurement between the ZHF sensor and
comparator device. \textgreater{}Information was extracted about how
repeated measurements were handled: (1) analysed each pair of data
separately; (2) treated each pair of data as independent; or (3) used
either analysis of variance or a random effects model as a way to
control for the dependent nature of the repeated measures data (Myles
and Cui 2007).

Using the revised Quality Assessment of Diagnostic Accuracy Studies
(QUADAS-2), each reviewer independently assessed the risk of bias for
the included studies (Whiting et al. 2011). Using guiding questions,
reviewers rated the risk of bias for patient selection, conduct of the
ZHF measurements, conduct of the comparator device measurements, and
timing and flow (eg, timing of ZHF and established core temperature
measurements, dropouts) as ``high'', ``low'' or ``unclear''. We worked
to minimize the risk of publication bias by conducting a comprehensive
search of multiple databases as well as an international clinical trial
registry (Glasziou et al. 2001). \textgreater{}Statistical analyses to
detect reporting bias were not conducted due to lack of validated
methods (Begg 2005). Although some meta-analyses of method comparison
studies have used tests for detecting funnel plot asymmetry (Niven et
al. 2015), simulations have revealed that such tests will result in
publication bias being incorrectly identified too often (Deeks,
Macaskill, and Irwig 2005).

In order to rate the quality of evidence, we applied the Grading Quality
of Evidence and Strength of Recommendation methodology (Schünemann et
al. 2008). Evidence was downgraded in accordance with study limitations,
inconsistency and imprecision. There were no circumstances in which
evidence was downgraded for indirectness as this systematic review only
included relevant studies. Although the possibility of publication bias
was not excluded, this bias was not formally assessed as it was not
considered sufficient enough to reason downgrading the quality of
evidence.

\hypertarget{data-synthesis-and-analysis}{%
\subsection{Data synthesis and
analysis}\label{data-synthesis-and-analysis}}

The objective for the meta-analysis was to estimate the population
limits of agreement between temperature measurements from the ZHF and
established comparator devices. \textgreater{}The framework for
meta-analysis of Bland-Altman method comparison studies based on limits
of agreement (LoA) approach was used (Tipton and Shuster 2017).

This method was selected because it parallels the approach used in
primary Bland-Altman studies, whereby an estimate is generated for the
pooled LoAs in the population (not just in the samples studied). The
`population LoA' is more broad than the LoA commonly reported in the
meta-analyses of Bland-Altman studies (Tipton and Shuster 2017). In
these instances, \textgreater{}the pooled LoAs are calculated using
\(\delta\pm2\sqrt{\sigma2+\tau2}\), where \(\delta\) is the average bias
across studies, \(\sigma2\) is the average within-study variation in
differences and \(\tau2\) is the variation in bias across studies.

Estimations of \(\delta\) and \(\sigma2\) were made using a weighted
least-squares model (similar to a random-effects approach), with the
associated SEs estimated using robust variance estimation (RVE). RVE was
used alternatively to model-based SEs as several studies included in the
systematic review used repeated-measures designs without accommodating
for the correlation between measurements (Hedges, Tipton, and Johnson
2010; Tanner-Smith, Tipton, and Polanin 2016; Tipton 2015). We used the
\textgreater{}method-of-moments estimator from ref 16 for the \(\tau2\)
parameter.

In accordance with ref 12, (1) measures of uncertainty were included
\textgreater{}when interpreting the LoA estimates by calculating the
outer \(95\%\) CIs for pooled LoA;

and (2) repeated measurements were adjusted if improperly adjusted for
in \textgreater{}individual studies (by using weights proportional to
the number of participants, not the total number of measurements).

The formulas for the above calculations from ref 12 can be found in
\href{}{file 1}. The R statistical program was used to conduct all
analyses (Team 2017). All data and R code (provided in ref 12) used in
the meta-analyses can be located \href{https://doi.org}{here}.

Prior to conducting the meta-analyses, the results from each study were
converted into a standard format, with bias meaning \(comparator-ZHF\)
measured in \(^\circ C\). In several studies, results were reported for
multiple groups of participants, therefore in the meta-analysis each of
these groups was treated as a separate `comparison'. Other studies
reported multiple sets of results, whereby analyses were conducted
between ZHF and various comparator devices used on the same participant.
These instances were also treated as a separate `comparison' if the
comparator devices were apart of separate meta-analyses groups. One
study reported intraoperative, postoperative and overall results for the
same participants. Only the paired measurements from the overall results
were included in the main and low risk bias analyses, leaving paired
measurements exclusively from the intraoperative and postoperative
timepoints to be included in respective meta-analyses subgroups. The
conventionally cited clinically acceptable agreement between ZHF and
comparator devices is 0.5 \(^\circ C\)
\((ref - most of the studies site this range from Eshragi...)\).
\textgreater{}It was deemed that outer confidence bounds for 95\% LoA
between ECT and ZHF (termed as `population limits of agreement') outside
of these bounds would

be clinically unacceptable. Sensitivity analysis for the primary
meta-analysis was performed based on risk of bias, whereby `unclear risk
of bias' was treated as `high risk' and `high risk of bias' studies were
excluded from analyses. As clinicians would be interested in the
accuracy of ZHF relative to the thermometer devices they use, and within
the clinical setting in which they use it we conducted subgroup analyses
according to the comparator device used (eg, core, peripheral,
nasopharangeal) and clinical setting (eg, intraoperative, intensuve care
unit).

\hypertarget{results}{%
\section{Results}\label{results}}

Sixteen studies were eligible for inclusion in this systematic review
(Figure 1).

\includegraphics[width=53.08in]{figure_one}

\begin{quote}
The characteristics of each study are in online supplementary
\href{link\%20to\%20file\%201}{file 1}.
\end{quote}

Using these studies, we conducted 6 different meta-analyses. The first
group compared ZHF to core (eg, arterial, bladder, esophageal, rectal)
temperature measurements and included 22 (73\%) comparisons from 14 (\%)
studies, and enrolled 607 participants with 179,821 paired measurements.
The second group compared ZHF to core temperature measurements in low
risk bias studies studies that scored a low risk of bias in all
categories) and included 10 (\%) comparisons from 5 (\%) studies, and
enrolled 273 participants with 104,294 paired measurements. The third
group compared ZHF to core temperature measurements in ICU patients and
included 7 (23\%) comparisons from 5 (31\%) studies, and enrolled 246
participants with 155,598 paired measurements. The fourth group compared
ZHF to core temperature measurement in intraoperative patients and
included 16 (53\%) comparisons from 10 (63\%) studies, and enrolled 464
participants with 24,223 paired measurements. The fifth group compared
ZHF to nasopharyngeal temperature measurements and included 4 (13\%)
comparisons from 4 (25\%) studies, and enrolled 297 participants with
109,819 paired measurements. The sixth group compared ZHF to sublingual
temperature measurement and included 2 (7\%) comparisons from 2 (13\%)
studies, and enrolled 107 participants with 22731 paired measurements.
The studies included in this systematic review evaluated a ZHF
temperature monitoring system manufactured by 3M. Previously known as
the SpotOn Temperature Monitoring System, the 3M ZHF device is now
referred to as the Bair Hugger Temperature Monitoring System. All
comparisons reported adherence to the ZHF device manufacturer
instructions and placed the sensor on the forehead of the participants.
High risk of bias was associated with patient selection for 13 (43\%)
comparisons from 9 (56\%) studies, conduct of ZHF and comparator
measurements in 10(33\%) comparisons from 8 (50\%) studies and 12 (40\%)
comparisons 9 (30\%) studies, respectively (mostly due to ZHF
measurements being taken with knowledge of the comparator measurements
and vice versa) and participant flow for 13 (43\%) studies. In 19 (50\%)
comparisons from 8 (50\%) studies, the authors had declared a conflict
of interest or receipt of funding or equipment from the manufacturer of
the ZHF device under evaluation.

\hypertarget{study-selection-and-description}{%
\subsection{Study selection and
description}\label{study-selection-and-description}}

\hypertarget{agreement-between-zhf-and-core-temperature-measurements}{%
\subsection{Agreement between ZHF and core temperature
measurements}\label{agreement-between-zhf-and-core-temperature-measurements}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-begg2005systematic}{}%
Begg, Colin B. 2005. ``Systematic Reviews of Diagnostic Accuracy Studies
Require Study by Study Examination: First for Heterogeneity, and Then
for Sources of Heterogeneity.'' \emph{Journal of Clinical Epidemiology}
58 (9). Elsevier Limited: 865.

\leavevmode\hypertarget{ref-deeks2005performance}{}%
Deeks, Jonathan J, Petra Macaskill, and Les Irwig. 2005. ``The
Performance of Tests of Publication Bias and Other Sample Size Effects
in Systematic Reviews of Diagnostic Test Accuracy Was Assessed.''
\emph{Journal of Clinical Epidemiology} 58 (9). Elsevier: 882--93.

\leavevmode\hypertarget{ref-frank1997perioperative}{}%
Frank, Steven M, Lee A Fleisher, Michael J Breslow, Michael S Higgins,
Krista F Olson, Susan Kelly, and Charles Beattie. 1997. ``Perioperative
Maintenance of Normothermia Reduces the Incidence of Morbid Cardiac
Events: A Randomized Clinical Trial.'' \emph{Jama} 277 (14). American
Medical Association: 1127--34.

\leavevmode\hypertarget{ref-glasziou2001systematic}{}%
Glasziou, Paul, Les Irwig, Chris Bain, and Graham Colditz. 2001.
\emph{Systematic Reviews in Health Care: A Practical Guide}. Cambridge
University Press.

\leavevmode\hypertarget{ref-hadian2006evidence}{}%
Hadian, Mehrnaz, and Michael R Pinsky. 2006. ``Evidence-Based Review of
the Use of the Pulmonary Artery Catheter: Impact Data and
Complications.'' \emph{Critical Care} 10 (3). BioMed Central: S8.

\leavevmode\hypertarget{ref-hedges2010robust}{}%
Hedges, Larry V, Elizabeth Tipton, and Matthew C Johnson. 2010. ``Robust
Variance Estimation in Meta-Regression with Dependent Effect Size
Estimates.'' \emph{Research Synthesis Methods} 1 (1). Wiley Online
Library: 39--65.

\leavevmode\hypertarget{ref-kurz1996perioperative}{}%
Kurz, Andrea, Daniel I Sessler, and Rainer Lenhardt. 1996.
``Perioperative Normothermia to Reduce the Incidence of Surgical-Wound
Infection and Shorten Hospitalization.'' \emph{New England Journal of
Medicine} 334 (19). Mass Medical Soc: 1209--16.

\leavevmode\hypertarget{ref-mann2012should}{}%
Mann, Rachel, and Simon M Gilbody. 2012. ``Should Methodological Filters
for Diagnostic Test Accuracy Studies Be Used in Systematic Reviews of
Psychometric Instruments? A Case Study Involving Screening for Postnatal
Depression.'' \emph{Systematic Reviews} 1 (1). BioMed Central: 9.

\leavevmode\hypertarget{ref-michelson1994reversible}{}%
Michelson, Alan D, Hollace MacGregor, Marc R Barnard, Anita S Kestin,
Michael J Rohrer, and C Robert Valeri. 1994. ``Reversible Inhibition of
Human Platelet Activation by Hypothermia in Vivo and in Vitro.''
\emph{Thrombosis and Haemostasis} 72 (05). Schattauer GmbH: 633--40.

\leavevmode\hypertarget{ref-myles2007using}{}%
Myles, PS, and James Cui. 2007. ``I. Using the Bland--Altman Method to
Measure Agreement with Repeated Measures.'' Oxford University Press.

\leavevmode\hypertarget{ref-niven2015accuracy}{}%
Niven, Daniel J, Jonathan E Gaudet, Kevin B Laupland, Kelly J Mrklas,
Derek J Roberts, and Henry Thomas Stelfox. 2015. ``Accuracy of
Peripheral Thermometers for Estimating Temperature: A Systematic Review
and Meta-Analysis.'' \emph{Annals of Internal Medicine} 163 (10).
American College of Physicians: 768--77.

\leavevmode\hypertarget{ref-preston2015improving}{}%
Preston, Louise, Christopher Carroll, Paolo Gardois, Suzy Paisley, and
Eva Kaltenthaler. 2015. ``Improving Search Efficiency for Systematic
Reviews of Diagnostic Test Accuracy: An Exploratory Study to Assess the
Viability of Limiting to Medline, Embase and Reference Checking.''
\emph{Systematic Reviews} 4 (1). BioMed Central: 82.

\leavevmode\hypertarget{ref-rohrer1992effect}{}%
Rohrer, Michael J, and ANITA M Natale. 1992. ``Effect of Hypothermia on
the Coagulation Cascade.'' \emph{Critical Care Medicine} 20 (10):
1402--5.

\leavevmode\hypertarget{ref-schunemann2008grading}{}%
Schünemann, Holger J, Andrew D Oxman, Jan Brozek, Paul Glasziou, Roman
Jaeschke, Gunn E Vist, John W Williams, et al. 2008. ``Grading Quality
of Evidence and Strength of Recommendations for Diagnostic Tests and
Strategies.'' \emph{Bmj} 336 (7653). British Medical Journal Publishing
Group: 1106--10.

\leavevmode\hypertarget{ref-tanner2016handling}{}%
Tanner-Smith, Emily E, Elizabeth Tipton, and Joshua R Polanin. 2016.
``Handling Complex Meta-Analytic Data Structures Using Robust Variance
Estimates: A Tutorial in R.'' \emph{Journal of Developmental and
Life-Course Criminology} 2 (1). Springer: 85--112.

\leavevmode\hypertarget{ref-team2017r}{}%
Team, R Core. 2017. ``R Core Team (2017). R: A Language and Environment
for Statistical Computing.'' \emph{R Found. Stat. Comput. Vienna,
Austria. URL Http://Www. R-Project. Org/., Page R Foundation for
Statistical Computing}.

\leavevmode\hypertarget{ref-tipton2015small}{}%
Tipton, Elizabeth. 2015. ``Small Sample Adjustments for Robust Variance
Estimation with Meta-Regression.'' \emph{Psychological Methods} 20 (3).
American Psychological Association: 375.

\leavevmode\hypertarget{ref-tipton2017framework}{}%
Tipton, Elizabeth, and Jonathan Shuster. 2017. ``A Framework for the
Meta-Analysis of Bland--Altman Studies Based on a Limits of Agreement
Approach.'' \emph{Statistics in Medicine} 36 (23). Wiley Online Library:
3621--35.

\leavevmode\hypertarget{ref-whiting2011quadas}{}%
Whiting, Penny F, Anne WS Rutjes, Marie E Westwood, Susan Mallett,
Jonathan J Deeks, Johannes B Reitsma, Mariska MG Leeflang, Jonathan AC
Sterne, and Patrick MM Bossuyt. 2011. ``QUADAS-2: A Revised Tool for the
Quality Assessment of Diagnostic Accuracy Studies.'' \emph{Annals of
Internal Medicine} 155 (8). Am Coll Physicians: 529--36.

\bibliographystyle{spbasic}
\bibliography{bibliography.bib}

\end{document}