### 2019-09-18 zhf_manuscript/zhf_manuscript.tex

% !TeX program = pdfLaTeX
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

%% load any required packages here

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\title{Accuracy and precision of zero-heat-flux temperature monitoring: a
systematic review and meta- analysis \thanks{Grants or other notes about the article that should go on the front page
should be placed here. General acknowledgments should be placed at the
end of the article.} }

    \titlerunning{Accuracy of zero-heat-flux temperature monitoring}

\author{  Aaron Conway \and  }

    \authorrunning{ Conway }

\institute{
        Aaron Conway \at
     Peter Munk Cardiac Centre, UHN \& Lawrence S. Bloomberg Faculty of
 Nursing, University of Toronto \\
     \email{\href{mailto:aaron.conway@utoronto.ca}{\nolinkurl{aaron.conway@utoronto.ca}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
The text of your abstract. 150 -- 250 words.
\\
\keywords{
        key \and
        dictionary \and
        word \and
    }

    \subclass{
                    MSC code 1 \and
                    MSC code 2 \and
            }

\end{abstract}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Core temperature monitoring is an important intervention within
perioperative and intensive care settings. Thermoregulatory dysfunction
is commonly associated with the induction of anesthesia and can lead to
adverse outcomes including cardiac dysrhythmias, altered hemostasis and
increased risk of surgical site infection (Frank et al. 1997; Kurz,
Sessler, and Lenhardt 1996; Michelson et al. 1994; Rohrer and Natale
1992). The pulmonary artery catheter is the reference standard for
continuous core temperature
monitoring\((ref - trouble finding a statement regarding this in the literature)\),
however, the invasive nature of this method renders an increased risk of
bloodstream infections and damage to surrounding tissue(Hadian and
Pinsky 2006). Although clinically accurate relative to pulmonary artery
temperature measurements, surrogate measures of core temperature at
esophageal, rectal, and bladder locations remain mildly invasive
interventions\((ref)\).

Zero-heat-flux thermometry is an alternative non-invasive method that
allows for continuous monitoring of core temperature. Originally
developed in the 1970s, zero-heat-flux technology was recently
implemented in the 3M SpotOn (3M, St Paul, MN), a single-use, disposable
sensor \((ref)\). In practice, the zero-heat-flux sensor, such as the 3M
SpotOn, is placed on the lateral surface of the forehead and initially
warmed to equilibrate the temperature of the skin surface to the
underlaying core tissues\((ref)\). Equipped with a thermal insulator,
the zero-heat-flux sensor eliminates heat loss to the environment to
allow for changes in core temperature to be directly reflected by a
change in skin surface temperature \((ref)\).

The agreement between zero-heat-flux thermometers and core as well as
other peripheral thermometers has been investigated in several studies
conducted within clinical settings. Appraisal of these studies and
synthesis of the results would aid clinicians in deciding the
appropriate circumstances in which zero-heat-flux thermometers may be
used. We aimed to determine if zero-heat-flux thermometers have
clinically acceptable accuracy and precision relative to established
core and peripheral temperature measurement devices. Accuracy is defined
as the average difference between temperature measurements from the
zero-heat-flux and comparator device and precision as the variance
(standard deviation) in the differences.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

A systematic review was conducted in accordance with a predetermined
protocol. The protocol was submitted to PROSPERO for registration but
due to delays in processing we had started data extraction by the time
it was reviewed. Therefore, the protocol did not meet the requirements
for registration. A copy of the submitted protocol can be accessed
\href{insert\%20link}{here}. The primary comparison for this review was
temperature measured from a zero-heat-flux thermometer versus a
temperature measured from a core site, which we defined as arterial,
esophaeal, bladder and rectal sites. Secondary comparisons were made
between zero-heat-flux thermometers and temperatures taken at peripheral
sites.

\hypertarget{inclusion-criteria}{%
\subsection{Inclusion criteria}\label{inclusion-criteria}}

Observational studies that reported temperature measurements from a
zero-heat-flux thermometer and comparator thermometer were included.
Studies involving a case control design were excluded due to potential
for overestimation of the intervention performance. Studies were
excluded if conducted on non-human subjects or outside of a clinical
healthcare setting. No publication date restrictions were
applied.Published conference abstracts were included if there was enough
information reported to appraise the quality of the study. There were no
language restrictions applied during the search.

\hypertarget{data-sources-and-searches}{%
\subsection{Data sources and searches}\label{data-sources-and-searches}}

Published studies were found by searching Medline and EMBASE from
January 2000 to July 2019. The Cochrane-recommended search strategy
combining terms for the `target condition' and `index test' was used
(Mann and Gilbody 2012). This search strategy is an efficient approach
for systematic reviews of diagnostic test accuracy studies.(Preston et
al. 2015) We also conducted forward citation searching, by using Google
Scholar to search the citations of the first article published on the
accuracy of zero-heat-flux thermometers. The search strategies used for
each data base can be located in \href{}{file 1}. Selection of studies
was undertaken independently by two reviewers using
\href{https://www.covidence.org}{Covidence}.

\hypertarget{data-extraction-and-quality-assessment}{%
\subsection{Data extraction and quality
assessment}\label{data-extraction-and-quality-assessment}}

Information was extracted regarding study characteristics (author, year
of publication, country, design, sample size, clinical setting, numbers
studied and analyses for each outcome), population characteristics
(inclusion and exclusion criteria) and temperature measurement
characteristics (placement of sensor, timing and methods of
measurements). The outcomes that were extracted included the mean bias
(eg, accuracy) and variance (eg, SD, precision) in temperature
measurement between the zero-heat-flux and comparator thermometers. We
also extracted information about how repeated measurements were handled.
In particular we assessed whether studies: (1) analysed each pair of
data separately; (2) treated each pair of data as independent; or (3)
used either analysis of variance or a random effects model as a way to
control for the dependent nature of the repeated measures data (Myles
and Cui 2007).

Two reviewers independently assessed the risk of bias for the included
studies using the revised Quality Assessment of Diagnostic Accuracy
Studies (QUADAS-2) (Whiting et al. 2011). Reviewers rated the risk of
bias for patient selection, conduct of the zero-heat-flux measurements,
conduct of the comparator thermometer measurements, and timing and flow
(eg, timing of ZHF and established core temperature measurements,
dropouts) as ``high'', ``low'' or ``unclear'' risk of bias. We worked to
minimize the risk of publication bias by conducting a comprehensive
search of multiple databases as well as an international clinical trial
registry (Glasziou et al. 2001). Statistical approaches for detection of
reporting bias were not conducted due to lack of validated methods (Begg
2005). ASimulations have revealed that tests for detecting funnel plot
asymmetry will result in publication bias being incorrectly identified
too often (Deeks, Macaskill, and Irwig 2005).

In order to rate the quality of evidence, we applied the Grading Quality
of Evidence and Strength of Recommendation methodology (Schünemann et
al. 2008). Evidence was downgraded in accordance with study limitations,
inconsistency and imprecision. There were no circumstances in which
evidence was downgraded for indirectness as this systematic review only
included relevant studies. Although the possibility of publication bias
was not excluded, this bias was not formally assessed as it was not
considered sufficient enough to reason downgrading the quality of
evidence.

\hypertarget{data-synthesis-and-analysis}{%
\subsection{Data synthesis and
analysis}\label{data-synthesis-and-analysis}}

The objective for the meta-analysis was to estimate the population
limits of agreement between temperature measurements from the ZHF and
established comparator thermometers. A framework for meta-analysis of
Bland-Altman method comparison studies developed by Tipton and Shuster
(2017) based on limits of agreement approach was used. This method was
selected because it parallels the approach used in primary Bland-Altman
studies, whereby an estimate is generated for the pooled LoAs in the
population (not just in the samples studied). The `population LoA' is
more broad than the LoA commonly reported in the meta-analyses of
Bland-Altman studies (Tipton and Shuster 2017). Pooled limits of
agreement were calculated using \(\delta\pm2\sqrt{\sigma2+\tau2}\),
where \(\delta\) is the average bias across studies, \(\sigma2\) is the
average within-study variation in differences and \(\tau2\) is the
variation in bias across studies.

Estimations of \(\delta\) and \(\sigma2\) were made using a weighted
least-squares model (similar to a random-effects approach), with the
associated standard errors estimated using robust variance estimation.
Robust variance estimation was used alternatively to model-based
standard errors as some studies included in the systematic review used
repeated-measures designs without accommodating for the correlation
between measurements (Hedges, Tipton, and Johnson 2010; Tanner-Smith,
Tipton, and Polanin 2016; Tipton 2015). We used the method-of-moments
estimator from DerSimonian and Laird (1986) for the \(\tau2\) parameter.

Measures of uncertainty were included in our meta-analyses by
calculating the outer \(95\%\) confidence intervals for pooled limits of
agreement. We also accounted for repeated measurements if they were not
properly adjusted for in individual studies. This was achieved by using
weights proportional to the number of participants, not the total number
of measurements. The R statistical program was used to conduct all
analyses (Team 2017). All data and R code used in the meta-analyses can
be located \href{}{here}.

Prior to conducting the meta-analyses, the results from each study were
converted into a standard format, with bias meaning
\(comparator thermometer-ZHF thermometer\) measured in \(^\circ C\). In
several studies, results were reported for multiple groups of
participants, therefore in the meta-analysis each of these groups was
treated as a separate `comparison'. Other studies reported multiple sets
of results, whereby analyses were conducted between ZHF and various
comparator devices used on the same participant. These instances were
also treated as a separate `comparison' if the comparator devices were
apart of separate meta-analyses groups. One study reported
intraoperative, postoperative and overall results for the same
participants. Only the paired measurements from the overall results were
included in the main and low risk bias analyses, leaving paired
measurements exclusively from the intraoperative and postoperative
timepoints to be included in respective meta-analyses subgroups.

The conventionally cited clinically acceptable agreement between ZHF and
comparator devices is 0.5 \(^\circ C\)
\((ref - most of the studies site this range from Eshragi...)\). It was
deemed that outer confidence bounds for 95\% LoA between zero-heat-flux
and core temperature measurements (termed as `population limits of
agreement') outside of these bounds wouldbe clinically unacceptable.

A sensitivity analysis for the primary comparison (ZHF versus
temperature measurement at core site) was performed based on risk of
bias, whereby `unclear risk of bias' was treated as `high risk' and
`high risk of bias' studies were excluded from analyses. As clinicians
would be interested in the accuracy of ZHF relative to the thermometer
devices they use, and within the clinical setting in which they use it,
we conducted subgroup analyses according to the comparator device used
(either core, sublingual or nasopharangeal) and clinical setting (either
intraoperative or intensive care unit).

\hypertarget{results}{%
\section{Results}\label{results}}

Fifteen studies were included in this systematic review (Figure 1). Two
studies reported only in abstract form were not included and assigned as
`studies awaiting classification' because there was insufficient
information provided.

\begin{figure}

{\centering \includegraphics[width=4.86in]{zhf_manuscript_files/figure-latex/fig-1-1}

}

\caption{PRISMA Flow Diagram}\label{fig:fig-1}
\end{figure}

\begin{quote}
The characteristics of each study are in online supplementary
\href{link\%20to\%20file\%201}{file 1}.
\end{quote}

The primary comparison of zero-heat-flux versus core temperature
measurements (eg, arterial, bladder, esophageal or rectal) consisted of
19 comparisons from 13 individual studies. In total, data from 576
participants with 179821 paired measurements were included in this
comparison (two studies did not report the total number of measurements
included in their analysis). The sensitivity analysis for the primary
comparison with only studies that were judged as low risk of bias across
all domains included 10 comparisons from 5 studies that enrolled 273
participants with 104294 paired measurements (two studies did not report
the total number of measurements included in their analysis). The third
group compared zero-heat-flux to core temperature measurements in ICU
patients and included 7 (23\%) comparisons from 5 (31\%) studies, and
enrolled 246 participants with 155,598 paired measurements. The fourth
group compared ZHF to core temperature measurement in intraoperative
patients and included 16 (53\%) comparisons from 10 (63\%) studies, and
enrolled 464 participants with 24,223 paired measurements. The fifth
group compared ZHF to nasopharyngeal temperature measurements and
included 4 (13\%) comparisons from 4 (25\%) studies, and enrolled 297
participants with 109,819 paired measurements. The sixth group compared
ZHF to sublingual temperature measurement and included 2 (7\%)
comparisons from 2 (13\%) studies, and enrolled 107 participants with
22731 paired measurements. The studies included in this systematic
review evaluated a ZHF temperature monitoring system manufactured by 3M.
Previously known as the SpotOn Temperature Monitoring System, the 3M ZHF
device is now referred to as the Bair Hugger Temperature Monitoring
System. All comparisons reported adherence to the ZHF device
manufacturer instructions and placed the sensor on the forehead of the
participants. High risk of bias was associated with patient selection
for 13 (43\%) comparisons from 9 (56\%) studies, conduct of ZHF and
comparator measurements in 10(33\%) comparisons from 8 (50\%) studies
and 12 (40\%) comparisons 9 (30\%) studies, respectively (mostly due to
ZHF measurements being taken with knowledge of the comparator
measurements and vice versa) and participant flow for 13 (43\%) studies.
In 19 (50\%) comparisons from 8 (50\%) studies, the authors had declared
a conflict of interest or receipt of funding or equipment from the
manufacturer of the ZHF device under evaluation.

\hypertarget{study-selection-and-description}{%
\subsection{Study selection and
description}\label{study-selection-and-description}}

\hypertarget{agreement-between-zhf-and-core-temperature-measurements}{%
\subsection{Agreement between ZHF and core temperature
measurements}\label{agreement-between-zhf-and-core-temperature-measurements}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\hypertarget{limitations}{%
\subsection{Limitations}\label{limitations}}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-begg2005systematic}{}%
Begg, Colin B. 2005. ``Systematic Reviews of Diagnostic Accuracy Studies
Require Study by Study Examination: First for Heterogeneity, and Then
for Sources of Heterogeneity.'' \emph{Journal of Clinical Epidemiology}
58 (9). Elsevier Limited: 865.

\leavevmode\hypertarget{ref-deeks2005performance}{}%
Deeks, Jonathan J, Petra Macaskill, and Les Irwig. 2005. ``The
Performance of Tests of Publication Bias and Other Sample Size Effects
in Systematic Reviews of Diagnostic Test Accuracy Was Assessed.''
\emph{Journal of Clinical Epidemiology} 58 (9). Elsevier: 882--93.

\leavevmode\hypertarget{ref-dersimonian1986meta}{}%
DerSimonian, Rebecca, and Nan Laird. 1986. ``Meta-Analysis in Clinical
Trials.'' \emph{Controlled Clinical Trials} 7 (3). Elsevier: 177--88.

\leavevmode\hypertarget{ref-frank1997perioperative}{}%
Frank, Steven M, Lee A Fleisher, Michael J Breslow, Michael S Higgins,
Krista F Olson, Susan Kelly, and Charles Beattie. 1997. ``Perioperative
Maintenance of Normothermia Reduces the Incidence of Morbid Cardiac
Events: A Randomized Clinical Trial.'' \emph{Jama} 277 (14). American
Medical Association: 1127--34.

\leavevmode\hypertarget{ref-glasziou2001systematic}{}%
Glasziou, Paul, Les Irwig, Chris Bain, and Graham Colditz. 2001.
\emph{Systematic Reviews in Health Care: A Practical Guide}. Cambridge
University Press.

\leavevmode\hypertarget{ref-hadian2006evidence}{}%
Hadian, Mehrnaz, and Michael R Pinsky. 2006. ``Evidence-Based Review of
the Use of the Pulmonary Artery Catheter: Impact Data and
Complications.'' \emph{Critical Care} 10 (3). BioMed Central: S8.

\leavevmode\hypertarget{ref-hedges2010robust}{}%
Hedges, Larry V, Elizabeth Tipton, and Matthew C Johnson. 2010. ``Robust
Variance Estimation in Meta-Regression with Dependent Effect Size
Estimates.'' \emph{Research Synthesis Methods} 1 (1). Wiley Online
Library: 39--65.

\leavevmode\hypertarget{ref-kurz1996perioperative}{}%
Kurz, Andrea, Daniel I Sessler, and Rainer Lenhardt. 1996.
``Perioperative Normothermia to Reduce the Incidence of Surgical-Wound
Infection and Shorten Hospitalization.'' \emph{New England Journal of
Medicine} 334 (19). Mass Medical Soc: 1209--16.

\leavevmode\hypertarget{ref-mann2012should}{}%
Mann, Rachel, and Simon M Gilbody. 2012. ``Should Methodological Filters
for Diagnostic Test Accuracy Studies Be Used in Systematic Reviews of
Psychometric Instruments? A Case Study Involving Screening for Postnatal
Depression.'' \emph{Systematic Reviews} 1 (1). BioMed Central: 9.

\leavevmode\hypertarget{ref-michelson1994reversible}{}%
Michelson, Alan D, Hollace MacGregor, Marc R Barnard, Anita S Kestin,
Michael J Rohrer, and C Robert Valeri. 1994. ``Reversible Inhibition of
Human Platelet Activation by Hypothermia in Vivo and in Vitro.''
\emph{Thrombosis and Haemostasis} 72 (05). Schattauer GmbH: 633--40.

\leavevmode\hypertarget{ref-myles2007using}{}%
Myles, PS, and James Cui. 2007. ``I. Using the Bland--Altman Method to
Measure Agreement with Repeated Measures.'' Oxford University Press.

\leavevmode\hypertarget{ref-preston2015improving}{}%
Preston, Louise, Christopher Carroll, Paolo Gardois, Suzy Paisley, and
Eva Kaltenthaler. 2015. ``Improving Search Efficiency for Systematic
Reviews of Diagnostic Test Accuracy: An Exploratory Study to Assess the
Viability of Limiting to Medline, Embase and Reference Checking.''
\emph{Systematic Reviews} 4 (1). BioMed Central: 82.

\leavevmode\hypertarget{ref-rohrer1992effect}{}%
Rohrer, Michael J, and ANITA M Natale. 1992. ``Effect of Hypothermia on
the Coagulation Cascade.'' \emph{Critical Care Medicine} 20 (10):
1402--5.

\leavevmode\hypertarget{ref-schunemann2008grading}{}%
Schünemann, Holger J, Andrew D Oxman, Jan Brozek, Paul Glasziou, Roman
Jaeschke, Gunn E Vist, John W Williams, et al. 2008. ``Grading Quality
of Evidence and Strength of Recommendations for Diagnostic Tests and
Strategies.'' \emph{Bmj} 336 (7653). British Medical Journal Publishing
Group: 1106--10.

\leavevmode\hypertarget{ref-tanner2016handling}{}%
Tanner-Smith, Emily E, Elizabeth Tipton, and Joshua R Polanin. 2016.
``Handling Complex Meta-Analytic Data Structures Using Robust Variance
Estimates: A Tutorial in R.'' \emph{Journal of Developmental and
Life-Course Criminology} 2 (1). Springer: 85--112.

\leavevmode\hypertarget{ref-team2017r}{}%
Team, R Core. 2017. ``R Core Team (2017). R: A Language and Environment
for Statistical Computing.'' \emph{R Found. Stat. Comput. Vienna,
Austria. URL Http://Www. R-Project. Org/., Page R Foundation for
Statistical Computing}.

\leavevmode\hypertarget{ref-tipton2015small}{}%
Tipton, Elizabeth. 2015. ``Small Sample Adjustments for Robust Variance
Estimation with Meta-Regression.'' \emph{Psychological Methods} 20 (3).
American Psychological Association: 375.

\leavevmode\hypertarget{ref-tipton2017framework}{}%
Tipton, Elizabeth, and Jonathan Shuster. 2017. ``A Framework for the
Meta-Analysis of Bland--Altman Studies Based on a Limits of Agreement
Approach.'' \emph{Statistics in Medicine} 36 (23). Wiley Online Library:
3621--35.

\leavevmode\hypertarget{ref-whiting2011quadas}{}%
Whiting, Penny F, Anne WS Rutjes, Marie E Westwood, Susan Mallett,
Jonathan J Deeks, Johannes B Reitsma, Mariska MG Leeflang, Jonathan AC
Sterne, and Patrick MM Bossuyt. 2011. ``QUADAS-2: A Revised Tool for the
Quality Assessment of Diagnostic Accuracy Studies.'' \emph{Annals of
Internal Medicine} 155 (8). Am Coll Physicians: 529--36.

\bibliographystyle{spbasic}
\bibliography{bibliography.bib}

\end{document}